---
title: "ST-512-HW5-Yefrid_Cordoba"
author: "Yefrid Cordoba"
format: html
editor: visual
---

## Problem 1

### a

Complete the ANOVA table

```{r}
a <- 5
dfa <- a-1
b <- 3
dfb <- b-1
dfab <- dfa * dfb
nT<- a * b * 6
dfe <- nT - (a * b)
SSA <- 375
SSB <- 462
SSAB <- 38
SST <- 1128
SSE <- SST - SSA - SSB - SSAB
MSA <- SSA / dfa
MSB <- SSB / dfb
MSAB <- SSAB / dfab
MSE <- SSE / dfe
F_A <- MSA / MSE
F_B <- MSB / MSE
F_AB <- MSAB / MSE
P_A <- pf(F_A, dfa, dfe, lower.tail = F)
P_B <- pf(F_B, dfb, dfe, lower.tail = F)
P_AB <- pf(F_AB, dfab, dfe, lower.tail = F)
```

| Source | df | SS | MS | F-value | P-value |
|------------|------------|------------|------------|------------|------------|
| **Factor A** | [`r dfa`]{.underline} | 375 | [`r MSA`]{.underline} | [`r round(F_A, 2)`]{.underline} | [`r P_A`]{.underline} |
| **Factor B** | [`r dfb`]{.underline} | 462 | [`r MSB`]{.underline} | [`r round(F_B, 2)`]{.underline} | [`r P_B`]{.underline} |
| **A:B** | [`r dfab`]{.underline} | 38 | [`r MSAB`]{.underline} | [`r round(F_AB, 2)`]{.underline} | [`r P_AB`]{.underline} |
| **Residual** | [`r dfe`]{.underline} | [`r SSE`]{.underline} | [`r round(MSE, 2)`]{.underline} |  |  |
| **Total** | [`r nT - 1`]{.underline} | 1128 |  |  |  |

### b

Assuming an $\alpha = 0.05$, and given that this is higher than the p-value for the interaction term (`r P_AB`) we conclude that there is no interaction between the two factors.

### c

Testing for levels of factor: since the p-value (`r P_A`)for the factor A is much smaller than the $\alpha = 0.05$, we conclude that factor A have an effect in the mean response.

### d

Testing for levels of factor B: since the p-value (`r P_B`)for the factor B is much smaller than the $\alpha = 0.05$, we conclude that factor B have an effect in the mean response.

## Problem 2

### a

Two way fixed model where interaction is not significant.

```{r}
a <- 4
b <- 3
n <- 3
nT <- a * b * n
SSE <- 720
dfa <- a - 1
dfb <- b - 1
dfe <- nT - (a * b)
MSE <- SSE / dfe
```

#### Tukey's HSD factor A

```{r}
SE <- sqrt(2*MSE/(b*n))
SE
qa <- qtukey(1-0.05, a, dfe)/sqrt(2)
qa
qHSDa <- SE * qa
qHSDa
```

#### Bonferroni factor A

```{r}
g = a*(a-1)/2
BA = qt(1-(0.05/(2*g)), nT - a * b) 
BA
BMSDa = SE * BA
BMSDa
```

### b

#### Tukey's HSD factor B

```{r}
SE <- sqrt(2*MSE/(a*n))
SE
qb <- qtukey(1-0.05, b, dfe)/sqrt(2)
qb
qHSDb <- SE * qb
qHSDb
```

#### Bonferroni factor A

```{r}
g = b*(b-1)/2
BB = qt(1-(0.05/(2*g)), nT - a * b) 
BB
BMSDb = SE * BB
BMSDb
```

## Problem 3

### a

Factor effects model $y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} +\epsilon_{ijk}$

Where:

$i = 1, 2, 3$ varieties\
$j = 1, 2, 3$ Temperatures\
$k = 1,2,..., 8$ Pots availables\
$y_{ijk} =$ the observation in the $i^{th}$ variety exposed to the $i^{th}$ temperature, $k^{th}$ pot.\
$\mu =$ a constant\
$\alpha_i =$ the effect of the $i^{th}$ variety.\
$\beta_j =$ the effect of the $j^{th}$ temperature.\
$(\alpha\beta)_{ij}$ interaction of the variety and temperature.\
$\epsilon_{ijk} =$ random error of the $k^{th}$ pot for the $i^{th}$ variety exposed to $j^{th}$ temperature.\

Assuming $\epsilon_{ijk} \sim N(0, \sigma^2)$, the errors are indepe

### b

2-way ANOVA partial output fo the model.

```{r}
dfm <- 8
SSm <- 71692.11
MSE <- 47.413
a <- 3
b <- 3
n <- 8
nT <- a * b * n
dfe<- nT - a * b
dfT <- dfm + dfe
SSE <- MSE * dfe
MSm <- SSm / dfm
F_m <- MSm / MSE
P_m <- pf(F_m, dfm, dfe, lower.tail = F)
```

| Source | df | Sum of squares | Mean square | F-value | P-value |
|------------|------------|------------|------------|------------|------------|
| Model | 8 | 71692.11 | [`r round(MSm, 2)`]{.underline} | [`r round(F_m, 2)`]{.underline} | \< 0.0001 |
| Error | [`r dfe`]{.underline} | [`r round(SSE, 2)`]{.underline} | 47.413 |  |  |
| Total | [`r dfT`]{.underline} |  |  |  |  |

```{r}
dfa <- a - 1
dfb <- b - 1
dfab <- dfa * dfb
MSA <- 1662.89
MSAB <- 1704.14
SSA <- MSA * dfa
SSAB <- MSAB *dfab
SSB <- SSm - SSA - SSAB
MSB <- SSB / dfb
F_A <- MSA / MSE
F_B <- MSB / MSE
F_AB <- MSAB / MSE
P_A <- pf(F_A, dfa, dfe, lower.tail = F)
P_B <- pf(F_B, dfb, dfe, lower.tail = F)
P_AB <- pf(F_AB, dfab, dfe, lower.tail = F)

```

| Source | df | Sum of squares | Mean square | F-value | P-value |
|------------|------------|------------|------------|------------|------------|
| Variety | [`r dfa`]{.underline} | [`r round(SSA, 2)`]{.underline} | 1662.89 | [`r round(F_A, 2)`]{.underline} | [`r P_A`]{.underline} |
| Temperature | [`r dfb`]{.underline} | [`r format(round(SSB, 2), scientific = F)`]{.underline} | [`r format(round(MSB, 2), scientific = F)`]{.underline} | [`r round(F_B, 2)`]{.underline} | [`r P_B`]{.underline} |
| Variety\*Temp | [`r dfab`]{.underline} | [`r round(SSAB, 2)`]{.underline} | 1704.14 | [`r round(F_AB, 2)`]{.underline} | [`r P_AB`]{.underline} |

```{r}
variety <- factor(rep(c(0, 6, 12), each = 3))
Temp <- factor(rep(c(15, 20, 25), times = 3))
Dry <- c(28.71, 53.57, 126.86, 40.14, 66.86, 103.57, 31.71, 53.43, 80.26)

data <- data.frame(variety, Temp, Dry)

```

### c

Interaction plot

```{r}
interaction.plot(Temp, variety, Dry)
```

The interaction plot shows that between the different treatments can be guessed that the interaction term is significant, since all of the lines are not following the same trend, one of them (variety P0)cross over the others.

### d

$H_o: (\alpha\beta)_{ij} = 0$ for all $i, j$\
$H_a: (\alpha\beta)_{ij} \neq 0$ for at least one.

Based on the P-value for the interaction term (`r P_AB`), and that we assumed an $\alpha = 0.05$, we reject the null hypothesis and have strong evidence to conclude that there is an interaction between variety of plant and temperature.

## Problem 4

```{r}
SSA <- 1255.24
SSAB <- 882.76
SST <- 4266.31
MSE <- 56.38
a <- 3
b <- 3
n <- 5
nT <- a * b * n
dfa <- a - 1
dfb <- b - 1
dfab <- dfa * dfb
dfe <- nT - a * b
SSE <- MSE * dfe
SSB <- SST - SSA - SSAB - SSE
MSA <- SSA / dfa
MSB <- SSB / dfb
MSAB <- SSA / dfab
F_A <- MSA / MSE
F_B <- MSB / MSE
F_AB <- MSAB / MSE
P_A <- pf(F_A, dfa, dfe, lower.tail = F)
P_B <- pf(F_B, dfb, dfe, lower.tail = F)
P_AB <- pf(F_AB, dfab, dfe, lower.tail = F)
```

| Source | df | Sum of squares | Mean square | F-value | P-value |
|--------------|------------|------------|------------|------------|------------|
| Climate (A) | [`r dfa`]{.underline} | 1255.24 | [`r MSA`]{.underline} | [`r round(F_A, 2)`]{.underline} | [`r P_A`]{.underline} |
| Body location (B) | [`r dfb`]{.underline} | [`r format(round(SSB, 2), scientific = F)`]{.underline} | [`r MSB`]{.underline} | [`r round(F_B, 2)`]{.underline} | [`r P_B`]{.underline} |
| Interaction (A\*B) | [`r dfab`]{.underline} | 882.76 | [`r MSA`]{.underline} | [`r round(F_AB, 2)`]{.underline} | [`r P_AB`]{.underline} |
| Error | [`r dfe`]{.underline} | [`r round(SSE, 2)`]{.underline} | 56.38 |  |  |
| Total | `r nT - 1` | 4266.31 |  |  |  |

I want to test whether there are significant differences between the number of bites in the arm or neck in cold and humid climates.

$H_o : \frac{\mu_{11} + \mu_{12}}{2} = \frac{\mu_{31} + \mu_{32}}{2}$

$L = \frac{1}{2}\mu_{11} + \frac{1}{2}\mu_{12} + (-\frac{1}{2})\mu_{31} = (-\frac{1}{2})\mu_{32}$

$\hat{L} = \frac{1}{2}\bar{y}_{11\cdot} + \frac{1}{2}\bar{y}_{12\cdot} + (-\frac{1}{2})\bar{y}_{31\cdot} + (-\frac{1}{2})\bar{y}_{32\cdot}$

```{r}
y11 <- 22.3
y12 <- 19.8
y31 <- 28
y32 <- 26
L_hat <- 0.5 * y11 + 0.5 * y12 - 0.5 * y31 - 0.5 * y32
L_hat
```

$\hat{L} =$ `r L_hat`

```{r}
n <- 5
s_L <- sqrt(MSE * ((4 * (1/2)^2)/n))
s_L
```

$s(\hat{L}) =$ `r s_L`

```{r}
t <- L_hat/s_L
t
pt(t, dfe)
```

Based on the P-values obtained (`r pt(t, dfe)`)which is lower than a significance level of $\alpha = 0.05$, we reject the null hypothesis and have strong evidence to conclude that there is a significant difference between the number of bites in the arm or neck in cold and humid climates.

## Problem 5

```{r}
n <- 72 + 180 + 15 + 32
SSE1 <- 42088
SSE2 <- 42122
SSE3 <- 42873
```

### a
Testing whether there is an interaction between gender and rank

$H_o: \beta_3 =0$\
$H_a: \beta_3 \neq 0$

Full model is model 1.\
Reduced model is model 2.

```{r}
dfr <- n - 3 
dff <- n - 4
F_a <- ((SSE2 - SSE1)/(dfr - dff))/(SSE1/dff)
P_a <- pf(F_a, dfr - dff, dff )
P_a
```

Given that the P-value = `r P_a` assessing whether the interaction term is significant, we fail to reject the null hypothesis, and have weak evidence to conclude that there is an interaction between gender and rank on ethics scores.

### b
Testing whether there is main effect for gender

$H_o: \beta_1 =0$\
$H_a: \beta_1 \neq 0$

Full model is model 1.\
Reduced model is model 3.

```{r}
dfr <- n - 3 
dff <- n - 4
F_b <- ((SSE3 - SSE1)/(dfr - dff))/(SSE1/dff)
P_b <- pf(F_b, dfr - dff, dff )
P_b
```

Given that the P-value = `r P_b` assessing whether the meain effect of gender is significant, we fail to reject the null hypothesis, and have weak evidence to conclude that gender has a main effect on ethics scores.