---
title: "ST-512-HW5-Yefrid_Cordoba"
author: "Yefrid Cordoba"
format: html
editor: visual
---

## Problem 1

### a

Complete the ANOVA table

```{r}
a <- 5
dfa <- a-1
b <- 3
dfb <- b-1
dfab <- dfa * dfb
nT<- a * b * 6
dfe <- nT - (a * b)
SSA <- 375
SSB <- 462
SSAB <- 38
SST <- 1128
SSE <- SST - SSA - SSB - SSAB
MSA <- SSA / dfa
MSB <- SSB / dfb
MSAB <- SSAB / dfab
MSE <- SSE / dfe
F_A <- MSA / MSE
F_B <- MSB / MSE
F_AB <- MSAB / MSE
P_A <- pf(F_A, dfa, dfe, lower.tail = F)
P_B <- pf(F_B, dfb, dfe, lower.tail = F)
P_AB <- pf(F_AB, dfab, dfe, lower.tail = F)
```

| Source | df | SS | MS | F-value | P-value |
|------------|------------|------------|------------|------------|------------|
| **Factor A** | [`r dfa`]{.underline} | 375 | [`r MSA`]{.underline} | [`r round(F_A, 2)`]{.underline} | [`r P_A`]{.underline} |
| **Factor B** | [`r dfb`]{.underline} | 462 | [`r MSB`]{.underline} | [`r round(F_B, 2)`]{.underline} | [`r P_B`]{.underline} |
| **A:B** | [`r dfab`]{.underline} | 38 | [`r MSAB`]{.underline} | [`r round(F_AB, 2)`]{.underline} | [`r P_AB`]{.underline} |
| **Residual** | [`r dfe`]{.underline} | [`r SSE`]{.underline} | [`r round(MSE, 2)`]{.underline} |  |  |
| **Total** | [`r nT - 1`]{.underline} | 1128 |  |  |  |

### b

Assuming an $\alpha = 0.05$, and given that this is higher than the p-value for the interaction term (`r P_AB`) we conclude that there is no interaction between the two factors.

### c

Testing for levels of factor: since the p-value (`r P_A`)for the factor A is much smaller than the $\alpha = 0.05$, we conclude that factor A have an effect in the mean response.

### d

Testing for levels of factor B: since the p-value (`r P_B`)for the factor B is much smaller than the $\alpha = 0.05$, we conclude that factor B have an effect in the mean response.

## Problem 2

### a

Two way fixed model where interaction is not significant.

```{r}
a <- 4
b <- 3
n <- 3
nT <- a * b * n
SSE <- 720
dfa <- a - 1
dfb <- b - 1
dfe <- nT - (a * b)
MSE <- SSE / dfe
```

#### Tukey's HSD factor A

```{r}
SE <- sqrt(2*MSE/(b*n))
SE
qa <- qtukey(1-0.05, a, dfe)/sqrt(2)
qa
MEa <- SE * qa
MEa
```
### b

#### Tukey's HSD factor B

```{r}
SE <- sqrt(2*MSE/(a*n))
SE
qb <- qtukey(1-0.05, b, dfe)/sqrt(2)
qb
MEb <- SE * qb
MEb
```

## Problem 3

### a

Factor effects model
$y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} +\epsilon_{ijk}$

Where:

$i = 1, 2, 3$ varieties\
$j = 1, 2, 3$ Temperatures\
$k = 1,2,..., 8$ Pots availables\
$y_{ijk} =$ the observation in the $i^{th}$ variety exposed to the $i^{th}$ temperature, $k^{th}$ pot.\
$\mu =$ a constant\
$\alpha_i =$ the effect of the $i^{th}$ variety.\
$\beta_j =$ the effect of the $j^{th}$ temperature.\
$(\alpha\beta)_{ij}$ interaction of the variety and temperature.\
$\epsilon_{ijk} =$ random error of the $k^{th}$ pot for the  $i^{th}$ variety exposed to $j^{th}$ temperature.\

Assuming $\epsilon_{ijk} \sim N(0, \sigma^2)$, the errors are indepe

```{r}
variety <- factor(rep(c(0, 6, 12), each = 3))
Temp <- factor(rep(c(15, 20, 25), times = 3))
Dry <- c(28.71, 53.57, 126.86, 40.14, 66.86, 103.57, 31.71, 53.43, 80.26)

data <- data.frame(variety, Temp, Dry)

```

### c
```{r}
interaction.plot(variety, Temp, Dry)
interaction.plot(Temp, variety, Dry)
```

